---
title: "Esercizio 3 Modelli Statistici"
author: "Potinga Marcelinio"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Parte 1

**Importing the txt file**
```{r import_data}
data <-  read.csv("https://raw.githubusercontent.com/marcel0501/Esercizi-Mod-Stat/refs/heads/main/ANTROP.TXT", sep="\t")
data$peso <- data$peso / 2.2 # Convert pounds to kg
```

**Stattistiche descrittive**
```{r descriptive_stats}
summary(data)
```

**Modelli di regressione aventi X=bicipite e Y=peso con le specificazioni lineare-lineare, log-lineare, log-log, lineare log e quadratica. **
```{r regression_models}
# Linear model
model_linear <- lm(peso ~ bicipite, data = data)
# Log-linear model
model_log_linear <- lm(log(peso) ~ bicipite, data = data)
# Log-log model
model_log_log <- lm(log(peso) ~ log(bicipite), data = data)
# Linear-log model
model_linear_log <- lm(peso ~ log(bicipite), data = data)
# Quadratic model
model_quadratic <- lm(peso ~ bicipite + I(bicipite^2), data = data)
# Best model selection based on F-statistic
models <- list(
  linear = model_linear,
  log_linear = model_log_linear,
  log_log = model_log_log,
  linear_log = model_linear_log,
  quadratic = model_quadratic
)
best_model <- NULL
best_f_stat <- -Inf
for (model in models) {
  f_stat <- summary(model)$fstatistic[1]
  if (f_stat > best_f_stat) {
    best_f_stat <- f_stat
    best_model <- model
  }
}
# Display the best model
summary(best_model)
# Plotting the best model
plot(data$bicipite, data$peso, main = "Best Model: Linear Log", xlab = "Bicipite", ylab = "Peso")
abline(best_model, col = "red")
# Residuals plot
plot(best_model$residuals, main = "Residuals of Best Model", ylab = "Residuals", xlab = "Index")
# Histogram of residuals
hist(best_model$residuals, main = "Histogram of Residuals", xlab = "Residuals", breaks = 20)
# QQ plot of residuals
qqnorm(best_model$residuals, main = "QQ Plot of Residuals")
qqline(best_model$residuals, col = "red")
#Residuals against fitted values
plot(best_model$fitted.values, best_model$residuals, main = "Residuals vs Fitted Values", xlab = "Fitted Values", ylab = "Residuals")
abline(h = 0, col = "red")
plot(best_model,which=2)
```

**Verifica eteroschedasticità per ogni modello**

```{r heteroscedasticity_check}
par(mfrow = c(1, length(models)))  # 1 row, N columns
# Plotting residuals vs fitted values for each model
for (i in seq_along(models)) {
  plot(fitted(models[[i]]), resid(models[[i]]),
       main = paste("Model", names(models)[i]),
       xlab = "Fitted", ylab = "Residuals")
  abline(h = 0, col = "red")
}
# Resetting the plotting layout 
par(mfrow = c(1, 1))
```
**Verifica normalità dei residui per ogni modello**
```{r normality_check}
par(mfrow = c(1, length(models)))  # 1 row, N columns
# QQ plot for each model
for (i in seq_along(models)) {
  qqnorm(resid(models[[i]]), main = paste("QQ Plot of Residuals -", names(models)[i]))
  qqline(resid(models[[i]]), col = "red")
}
# Resetting the plotting layout
par(mfrow = c(1, 1))
```


##Parte 2
**Importing the dataset**
```{r import_data_part2}
data2 <- read.csv("https://raw.githubusercontent.com/marcel0501/Esercizi-Mod-Stat/refs/heads/main/nazioni.csv", sep=";")
```

```{r scatter_plot}
numeric <- unlist(lapply(data2, is.numeric))
library(psych)
library(corrplot)
# Scatter plot all the variables
pairs.panels(data2[,numeric], lm=T)
par(mfrow=c(1,1))

Corr <- round(cor(data2[,numeric]), 3)
corrplot(Corr, method="number")

summary(data2)

freq_table <- table(data2$relig)
barplot(freq_table, main="Distribuzione delle Religioni", xlab="Religione", ylab="Frequenza", col=rainbow(length(freq_table)))
```
**Regression on PIL with density, urban, maleexp, femexp, religion**
```{r regression_pil}
table(data2$relig)
data2$relig <- factor(data2$relig)
data2$relig
model0 <- lm(pil ~ densita + urbana + vitamas + vitafem + relig, data = data2)
summary(model0)
require(moldel0)
install.packages("SparseM")

```
